---
execute: 
  echo: false
  message: false
  warning: false
---

# Case studies {#sec-case-studies}

```{r}
#| echo: false
library(knitr)
knit_print.gt <- function(x, ...) {
  # Two steps to avoid most Quarto changes of my table styles: 
  # 1. as_raw_html() to use table styles *inline*
  # 2. wrap output in a div that resets all Quarto styles
  stringr::str_c(
    "<div style='all:initial';>\n", 
    gt::as_raw_html(x), 
    "\n</div>"
  ) |> 
    knitr::asis_output()
}
registerS3method(
  "knit_print", 'gt_tbl', knit_print.gt, 
  envir = asNamespace("gt") 
  # important to overwrite {gt}s knit_print
)
```

## Fitness pricing

In this chapter we're going to build this Fitness pricing table I found [online](https://codepen.io/supacode/pen/ExjbmgG).
It's not really a data table but it's a fun exercise to build this.

```{r}
library(tidyverse)
library(gt)
levels <- c('Standard', 'Popular', 'Golden', 'Ultimate')
prices <- c(15, 25, 35, 50)
names(prices) <- levels

features <- c('Beginner Classes', 'Training Overview', 'Personal Training', 'Olympic Weightlifting', 'Foundation Training')

dat <- tibble(
  level = levels,
  monthly_price = prices[levels],
  features = list(features[1], features[1:2], features[1:3], features),
  booking_link = 'https://www.youtube.com/watch?v=dQw4w9WgXcQ'
)

level_colors <- c(
  Standard = "#c40d53",
  Popular = "#26559b",
  Golden = "#f90",
  Ultimate =  "#0d833e"
)

```


```{r}
#| column: screen
unnested_dat <- dat |>
  unnest(features) |>
  select(level, features)

rearranged_data <- unnested_dat |>
  mutate(in_level = TRUE)  |>
  complete(
    expand(unnested_dat, level, features),
    fill = list(in_level = FALSE)
  ) |>
  mutate(
    id = paste(level, features, sep = '_')
  ) |>
  pivot_wider(
    id_cols = c(id, in_level),
    names_from = level,
    values_from = features
  ) |>
  select(-id) |>
  mutate(across(.cols = -in_level, .fns = ~factor(., levels = features)))


fontawesome_checkmark <- function(color) {
  glue::glue(
    '<svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 512 512" fill="{color}"><!--! Font Awesome Pro 6.2.0 by @fontawesome - https://fontawesome.com License - https://fontawesome.com/license (Commercial License) Copyright 2022 Fonticons, Inc. --><path d="M470.6 105.4c12.5 12.5 12.5 32.8 0 45.3l-256 256c-12.5 12.5-32.8 12.5-45.3 0l-128-128c-12.5-12.5-12.5-32.8 0-45.3s32.8-12.5 45.3 0L192 338.7 425.4 105.4c12.5-12.5 32.8-12.5 45.3 0z"/></svg>'
  )
}

fontawesome_xmark <- function(color) {
  glue::glue('<svg xmlns="http://www.w3.org/2000/svg" viewBox="0 0 320 512" fill="{color}" width="75%" height="75%"><!--! Font Awesome Pro 6.2.0 by @fontawesome - https://fontawesome.com License - https://fontawesome.com/license (Commercial License) Copyright 2022 Fonticons, Inc. --><path d="M310.6 150.6c12.5-12.5 12.5-32.8 0-45.3s-32.8-12.5-45.3 0L160 210.7 54.6 105.4c-12.5-12.5-32.8-12.5-45.3 0s-12.5 32.8 0 45.3L114.7 256 9.4 361.4c-12.5 12.5-12.5 32.8 0 45.3s32.8 12.5 45.3 0L160 301.3 265.4 406.6c12.5 12.5 32.8 12.5 45.3 0s12.5-32.8 0-45.3L205.3 256 310.6 150.6z"/></svg>')
}

feature_tables <- character(length(levels))
for (level in levels) {

  #level <- 'Ultimate'
  level_data <- rearranged_data |>
    select(in_level, !!level) |>
    filter(!is.na(!!sym(level))) |>
    arrange(!!sym(level))

  tbl <- level_data |>
  mutate(
    across(everything(), as.character),
    in_level = ifelse(
      parse_logical(in_level),
      fontawesome_checkmark(level_colors[level]),
      fontawesome_xmark(level_colors[level])
    )
  ) |>
  gt() |>
  fmt_markdown(columns = 'in_level') |>
  cols_width(in_level ~ px(25), everything() ~ px(175)) |>
  tab_options(
    column_labels.hidden = TRUE,
    table.font.names = 'Source Sans Pro',
    table_body.border.top.style = 'none',
    table.border.top.style = 'none',
    table_body.border.bottom.style = 'none',
    table.border.bottom.style = 'none',
    table_body.hlines.style = 'none',
  ) |>
    as_raw_html()

  feature_tables[which(level == levels)] <- tbl
}

tables_tib <- tibble(
  level = levels,
  table = feature_tables
) |>
  pivot_wider(names_from = level, values_from = table)

price_tib <- tibble(
  level = levels,
  price = paste0('$<span style="font-size:40px;">', prices, '</span>/month')
) |>
  pivot_wider(names_from = level, values_from = price)


style_links <- function(link, color) {
  htmltools::a(
      href = link,
      "Book now",
      style = glue::glue("border-radius: 5px;color: white;background-color: {color};border-radius: 5px;padding: 8px 20px;display: inline-block;text-decoration:none")
  )
}

links_tib <- tibble(
  level = levels,
  level_colors = level_colors[level],
  booking_link = 'https://www.youtube.com/watch?v=dQw4w9WgXcQ'
) |>
  mutate(
    booking_link = map2(booking_link, level_colors, style_links),
    booking_link = map_chr(booking_link, ~as.character(.x))
  ) |>
  select(-level_colors) |>
  pivot_wider(names_from = level, values_from = booking_link)

levels_text_size <- px(30)

bind_rows(price_tib, tables_tib, links_tib) |>
  mutate(dummy1 = '', .after = 1) |>
  mutate(dummy2 = '', .after = 3) |>
  mutate(dummy3 = '', .after = 5) |>
  gt(id = 'my_tbl') |>
  fmt_markdown(columns = everything()) |>
  cols_align(align = 'center') |>
  cols_label(
    dummy1 = '',
    dummy2 = '',
    dummy3 = ''
  ) |>
  cols_width(
    dummy1 ~ px(15),
    dummy2 ~ px(15),
    dummy3 ~ px(15)
  ) |>
  tab_style(
    style = list(
      cell_fill(color = level_colors['Standard']),
      cell_text(color = 'white', weight = 'bold')
    ),
    locations = list(
      cells_body(columns = 'Standard', rows = 1),
      cells_column_labels(column = 'Standard')
    )
  ) |>
  tab_style(
    style = list(
      cell_fill(color = level_colors['Popular']),
      cell_text(color = 'white', weight = 'bold')
    ),
    locations = list(
      cells_body(columns = 'Popular', rows = 1),
      cells_column_labels(column = 'Popular')
    )
  ) |>
  tab_style(
    style = list(
      cell_fill(color = level_colors['Golden']),
      cell_text(color = 'white', weight = 'bold')
    ),
    locations = list(
      cells_body(columns = 'Golden', rows = 1),
      cells_column_labels(column = 'Golden')
    )
  ) |>
  tab_style(
    style = list(
      cell_fill(color = level_colors['Ultimate']),
      cell_text(color = 'white', weight = 'bold')
    ),
    locations = list(
      cells_body(columns = 'Ultimate', rows = 1),
      cells_column_labels(column = 'Ultimate')
    )
  ) |>
  tab_style(
    style = cell_text(size = levels_text_size),
    locations = cells_column_labels()
  ) |>
  tab_style(
    style = list(cell_fill(color = 'white')),
    locations = list(
      cells_body(columns = contains('dummy')),
      cells_column_labels(columns = contains('dummy'))
    )
  ) |>
  tab_style(
    style = cell_borders(sides = 'bottom', color = '#D3D3D3'),
    locations = cells_body(rows = 3, columns = c(1, 3, 5, 7))
  ) |>
  tab_options(
    table_body.border.top.style = 'none',
    table.border.top.style = 'none',
    table_body.border.bottom.style = 'none',
    table.border.bottom.style = 'none',
    table_body.hlines.style = 'none',
    table_body.vlines.style = 'solid',
    column_labels.border.top.style = 'none',
    column_labels.border.bottom.style = 'none',
    column_labels.border.lr.style = 'solid', # not working, set in css
    column_labels.border.lr.width = px(1),
    column_labels.padding = px(1),
    data_row.padding = px(2),
    table.font.names = 'Source Sans Pro',
    heading.title.font.size = px(45),
    heading.padding = px(10),
    heading.border.bottom.style = 'none'
  ) |>
  tab_header(
    title = 'Fitness Pricing Table'
  ) |>
  tab_footnote(
    footnote = 'Design/Inspiration: @supacode | {gt} Remake: @rappa753',
    placement = 'right'
  ) |>
  opt_css(
    '.gt_footnote {
      text-align: right; padding-top: 5px;
    }

    .gt_title {
      font-family:"Oleo Script";
    }

    #my_tbl .gt_col_heading {
      border-left-style:solid;
      border-right-style:solid;
    }
    
    #my_tbl thead, tbody, tfoot, tr, td, th {
      border-color: inherit;
      border-style: solid;
      border-width: 0;
    }
    
    
    #my_tbl a {
    &:hover {
        transform: translateY(-3px);
        box-shadow: 0 5px 5px rgba(0, 0, 0, 0.4);
      }
      &:active {
        box-shadow: inset 0 -3px 5px rgba(0, 0, 0, 0.4);
      }
    }



    '
  )

```

## NYT bestseller

This one is a recreation of an awesome table [Tanya Shapiro](https://twitter.com/tanya_shapiro/status/1584616721251725312) made with `{ggplot2}`.
It is a huge table.
So you'll probably need to look at this on a large screen.
But just to be safe.
Here's a [screenshot of the table](img/screenshot_table.png) as well.


```{r}
library(tidyverse)
library(gt)
library(gtExtras)
```


```{r}
nyt_dat <- readr::read_tsv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-05-10/nyt_full.tsv') |> 
  mutate(decade = (year %/% 10) * 10) 
```


```{r}
top5_authors_by_decade <- nyt_dat |> 
  filter(between(year, 1960, 2019)) |> 
  count(author, decade, sort = T, name = 'weeks') |> 
  group_by(decade) |> 
  slice_max(weeks, n = 5) |> 
  ungroup()
```


```{r}
find_best_book <- function(decade, author) {
  best_books <- nyt_dat |> 
    filter(decade == !!decade, author == !!author) |> 
    count(title, sort = TRUE)
  
  best_books[[1, 'title']] |> str_to_title()
}

find_number_of_weeks_per_year <- function(decade, author){
  nyt_dat |> 
    filter(decade == !!decade, author == !!author) |> 
    count(year, name = 'weeks') |> 
    complete(tibble(year = decade:(decade + 9)), fill = list(weeks = 0)) |> 
    arrange(year)  |> 
    pull(weeks)
}


top5_authors_and_book_by_decade <- top5_authors_by_decade |> 
  mutate(
    best_book = map2_chr(decade, author, find_best_book),
    sparkline_weeks = map2(decade, author, find_number_of_weeks_per_year)
  )
```


```{r}
image_links <- tibble(
  author = top5_authors_and_book_by_decade |> pull(author) |> unique(),
  img = c(
    'https://images.gr-assets.com/authors/1327446818p8/77616.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/c/c5/John_O%27Hara_cph.3b08576.jpg/1024px-John_O%27Hara_cph.3b08576.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/8/88/Taylor_caldwell_a.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/0/0c/Irving_Wallace%2C_1972.jpg/330px-Irving_Wallace%2C_1972.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/d/d6/Leon_Uris_%28cropped%29.jpg',
    'https://images-na.ssl-images-amazon.com/images/I/41RMdx8BJHL.__01_SX120_CR0,0,120,120__.jpg',
    'https://upload.wikimedia.org/wikipedia/en/a/a2/Robert_Ludlum_%281927-2001%29.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/4/49/Herman_Wouk_%28cropped%29.jpg/330px-Herman_Wouk_%28cropped%29.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/c/cf/Agatha_Christie.png/330px-Agatha_Christie.png',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/9/9d/Frederick_Forsyth_-_01.jpg/375px-Frederick_Forsyth_-_01.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/e/e3/Stephen_King%2C_Comicon.jpg/330px-Stephen_King%2C_Comicon.jpg',
    'https://images1.penguinrandomhouse.com/author/29599',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/d/dc/James_Albert_Michener_%C2%B7_DN-SC-92-05368.JPEG/330px-James_Albert_Michener_%C2%B7_DN-SC-92-05368.JPEG',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/9/98/Tom_Clancy_at_Burns_Library_cropped.jpg/330px-Tom_Clancy_at_Burns_Library_cropped.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/5/5f/Grisham_John_by_C_Harrison_.jpg/300px-Grisham_John_by_C_Harrison_.jpg',
    'https://upload.wikimedia.org/wikipedia/en/7/70/Robert_James_Waller.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/b/bf/Mary_Higgins_Clark_at_the_Mazza_Museum.jpg/330px-Mary_Higgins_Clark_at_the_Mazza_Museum.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/8/8b/Dan_Brown_bookjacket_cropped.jpg/330px-Dan_Brown_bookjacket_cropped.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/1/1d/James_Patterson.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/a/af/Nicholas-Sparks-Autograph-1-4-06.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/3/3d/David_Baldacci_-_2015_National_Book_Festival_%286%29.jpg/330px-David_Baldacci_-_2015_National_Book_Festival_%286%29.jpg',
    'https://upload.wikimedia.org/wikipedia/commons/thumb/6/61/Anthony_Doerr_%282015%29.jpg/330px-Anthony_Doerr_%282015%29.jpg'
  )
)

full_dat <- top5_authors_and_book_by_decade |> 
  left_join(image_links)

```




```{r}

map2color<-function(x, pal = rev(RColorBrewer::brewer.pal(11, 'Spectral')), limits=NULL){
    if(is.null(limits)) limits <- range(x)
    pal[findInterval(x,seq(limits[1],limits[2],length.out=length(pal)+1), all.inside=TRUE)]
}

create_point_div <- function(color, size) {
  glue::glue(
    '<span style="height: {size}em;width: {size}em;background-color: {color};border-radius: 50%;margin-top:4px;display:inline-block;margin-left:2px;"></span>'
  )
}

format_text <- function(weeks, author, best_book, colors) {
  glue::glue(
    '<span style = "color:white;font-weight:lighter;font-size:12pt;">{str_to_upper(weeks)} WEEKS ON THE LIST</span> {create_point_div(colors, 0.75)}',
    '<br>',
    '<span style = "color:white;font-weight:bold;font-size:22pt;">{author}</span>',
    '<br>',
    '<span style = "color:white;font-size:12pt;">{best_book}</span>'
  )
}
```



```{r}
# can be higher than 52 bc of multiple books
highest_week_count_per_year <- full_dat |> 
  pull(sparkline_weeks) |> 
  map_dbl(max) |>
  max()
create_sparkline <- function(sparkline) {
  ggplot() +
    geom_line(
      mapping = aes(x = seq_along(sparkline), y = sparkline),
      color = 'white',
      size = 0.5
    ) +
    annotate(
      'text',
      x = c(1, 10),
      y = sparkline[c(1, 10)],
      label = sparkline[c(1, 10)],
      color = 'white',
      hjust = c(1.2, -0.2),
      size = 2.5
    ) +
    coord_cartesian(xlim = c(-5, 15), ylim = c(0, highest_week_count_per_year)) +
    theme_void() +
    theme(plot.background = element_rect(fill = 'black'))
}


for (decade in seq(1960, 2010, 10)) {
  decade_data <- full_dat |> filter(decade == !!decade)
  
  for (k in 1:5) {
    p <- decade_data |> 
      pluck('sparkline_weeks', k) |> 
      create_sparkline()
    
    ggsave(
      glue::glue('img/{decade}_{k}.png'),
      plot = p,
      height = 200,
      width = 400,
      unit = 'px'
    )
  }
  
}
```


```{r}

create_decade_table <- function(decade, img_size = 75) {
  full_dat |> 
    mutate(color = map2color(weeks)) |> 
    filter(decade == !!decade) |> 
    mutate(
      author = if_else(author == 'Robert James Waller', 'Robert Waller', author),
      rank = 1:5,
      sparkline_weeks = paste0(
        here::here('img/'),
        decade,
        '_',
        rank,
        '.png'
      ),
      decade = paste0(decade, 's'),
      test = pmap(.l = list(weeks, author, best_book, color), format_text)
  ) |> 
  select(
    rank, decade, test, sparkline_weeks, img
  ) |> 
  gt(groupname_col = 'decade') |> 
  gt_img_circle(column = 'img', height = img_size) |> 
  gt_img_rows(columns = 'sparkline_weeks', img_source = 'local', height = 75) |> 
  cols_width(
    img ~ px(82), # For some reason it needs to be a little more than 75px
    test ~ px(300),
    sparkline_weeks ~ px(100)
  ) |> 
  cols_align('left', columns = 'test') |> 
  fmt_markdown(columns = c('test')) |> 
  tab_options(
    column_labels.hidden = TRUE,
    table.background.color = 'black',
    table.font.color = 'white',
    table.font.names = 'Open Sans',
    data_row.padding = px(0)
  ) |> 
  tab_style(
    style = cell_text(size = '15pt', weight = 'bold', v_align = 'top'),
    locations = cells_body('rank')
  ) |> 
  opt_css(
    '.gt_group_heading_row {
      text-align:center;
    }
    '
  )
}
```


```{r}
#| column: screen
# Can't use inline css because it for some reason destroys the first column of each
# table otherwise
raw_tables <- map(seq(1960, 2010, 10), create_decade_table) |> 
  map_chr(as_raw_html, inline_css = FALSE)

## Tanya, in case you're reading this: Thank you for coding the title.
## I mean, thank you for the whole cool table. 
## But especially thank you for coding the title.
## I didn't want to piggyback your code as much as possible but after working for hours
## to get the {gt} mechanics right, I just couldn't handly having to `<span>` the texts.
## Your code was a true life saver here.
title <- paste0(
  "<span style='font-family:Chomsky;font-size:42pt;color:white;'> The New York Times</span>",
  "<span style='font-family:opensans;font-size:24pt;color:white;'> **Best Selling Authors**</span>",
  "<br><span style='font-family:opensans;font-size:18pt;color:#D6D6D6'>Top authors by decade. Ranking based on number of weeks author appeared on list. Sparkline depicts total weeks<br>by year (counts multiple books). Top performing book included beneath each author's name. Data from Post45 Data Collective.</span><br>"
)

tbl <- tibble(
  col1 = raw_tables[c(1, 4)],
  col2 = raw_tables[c(2, 5)],
  col3 = raw_tables[c(3, 6)]
) |> 
  gt(id = 'my_tbl') |> 
  fmt_markdown(columns = everything()) |> 
  tab_options(
    table.background.color = 'black',
    table.font.color = 'white',
    table.font.names = 'Open Sans',
    column_labels.hidden = TRUE,
    row_group.border.top.style = 'none',
    row_group.border.bottom.style = 'none',
    table.border.bottom.style = 'solid',
    table.border.bottom.width = px(1),
    table.border.top.style = 'none',
    table_body.border.bottom.style = 'none',
    table_body.border.top.style = 'none',
    heading.border.bottom.style = 'none',
    column_labels.border.top.style = 'none',
  ) |> 
  opt_css(
    '.gt_group_heading_row {
      text-align:center;
      font-size:16pt;
    }
    
    #my_tbl .gt_footnote {
      text-align: right; 
      padding-top: 20px;
      padding-bottom:5px;
      font-family:"Open Sans";
      font-size:10pt;
      font-weight:bold;
    }
    
    #my_tbl .gt_row {
      border-top-color: grey;
      border-bottom-color: grey;
    }
    
    .gt_table_body {line-height:1.2;}
    
    '
  ) |> 
  tab_header(
    title = md(title)
  ) |> 
  tab_style(
    locations = cells_body(rows = 1),
    style = cell_borders(style = 'hidden')
  ) |> 
  tab_footnote(
    html(glue::glue(
      '{ggplot2} Original: <<fontawesome::fa("twitter")>>@tanya_shapiro | 
      {gt} recreation: <<fontawesome::fa("twitter")>>@rappa753',
      .open = "<<", .close = ">>"
    ))
  ) |> 
  cols_width(
    col1 ~ px(600),
    col2 ~ px(600),
    col3 ~ px(600)
  )


tbl 
```


## LaTeX formulas

In the absence of current support for formulas with MathJax

```{r}
#| include: false
#| eval: false
exams::tex2image(
  '$\\overline{X_n} = \\sum\\limits_{k = 1}^{n} x_k $', 
  format = 'svg', 
  dir = here::here(), 
  name = 'formula'
)
svg_formula_black <- readLines('formula.svg') |> 
  str_flatten() 
  
svg_formula_white <- svg_formula_black |> 
  str_replace_all('stroke:rgb\\(0%,0%,0%\\)', 'stroke:#FFFFFF') |> 
  str_replace_all('fill:rgb\\(0%,0%,0%\\)', 'fill:#FFFFFF')

exams::tex2image(
  '$\\int_0^\\infty f(x) dx$', 
  format = 'svg', 
  dir = here::here(), 
  name = 'formula2',
  packages = 'amsmath'
)
svg_formula_black <- readLines('formula2.svg') |> 
  str_flatten() |> 
  str_replace_all('glyph', 'bam')

tibble(char = svg_formula_black, char2 = 'bla') |> 
  gt() |> 
  opt_stylize(style = 3) |> 
  tab_spanner(
    label = md(svg_formula_white),
    columns = 1:2
  ) |> 
  fmt_markdown(columns = 'char') |> 
  tab_header(title = 'This is a table with formulas as svgs')
```


